---
title: "diabetes prediction"
output: html_document
date: "2024-05-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Load the package

```{r}
library(naniar)
library(ggplot2)
library(tidyverse)
library(tidymodels)
library(corrplot)
library(ggthemes)
library(finalfit)
library(conflicted)

tidymodels_prefer()
```

#### Data Preparation

```{r}
diabete_data = read_csv("diabetes_prediction_dataset.csv", show_col_types = F)
head(diabete_data)
```

##### tidy the dataset

```{r}
dim(diabete_data)

diabete_data$gender = as.factor(diabete_data$gender)
diabete_data$hypertension = as.factor(diabete_data$hypertension)
diabete_data$heart_disease = as.factor(diabete_data$heart_disease)
diabete_data$smoking_history = as.factor(diabete_data$smoking_history)
diabete_data$diabetes = as.factor(diabete_data$diabetes)
```

##### dataset summary

```{r}
summary(diabete_data)
```

##### missing data

```{r}
diabete_data %>%
  missing_plot()
```

#### EDA

##### Distribution of cltv

```{r}
diabete_data %>%
  ggplot(aes(diabetes)) +
  geom_bar(width = 0.4, aes(fill = diabetes)) +
  scale_fill_manual(values = c("#1f77b4", "#ff7f0e"), 
                    labels = c("Non-diabetic", "Diabetic")) +
  theme_gray()
```

##### Correlation of numeric variables and factor variable with 2 levels

```{r}
diabete_data %>%
  select(where(~is.numeric(.) || (is.factor(.) && nlevels(.) == 2))) %>%
  mutate(across(where(~is.factor(.) && nlevels(.) == 2), as.numeric)) %>%
  cor() %>%
  corrplot(method = "circle",
           diag = F,
           tl.col = "black",
           addCoef.col = 1, 
           number.cex = 0.7)
```

##### diabetes v.s. age and BMI

```{r}
p1 = diabete_data %>%
  ggplot(aes(x = diabetes, y = age, fill = diabetes)) +
  geom_boxplot() +
  scale_fill_manual(values = c("#1f77b4", "#ff7f0e"), 
                    labels = c("Non-diabetic", "Diabetic")) +
  theme_gray()

p2 = diabete_data %>%
  ggplot(aes(x = diabetes, y = bmi, fill = diabetes)) +
  geom_boxplot() +
  scale_fill_manual(values = c("#1f77b4", "#ff7f0e"), 
                    labels = c("Non-diabetic", "Diabetic")) +
  theme_gray()

grid.arrange(p1, p2, ncol = 2)
```

##### age v.s. BMI

```{r, message = FALSE}
diabete_data %>%
  ggplot(aes(x = age, y = bmi)) +
  geom_point() +
  geom_smooth() +
  theme_gray()
```

##### diabetes v.s. blood glucose level and HbA1c level

```{r}
ppp1 = diabete_data %>%
  ggplot(aes(x = diabetes, y = blood_glucose_level, fill = diabetes)) +
  geom_boxplot() +
  scale_fill_manual(values = c("#1f77b4", "#ff7f0e"), 
                    labels = c("Non-diabetic", "Diabetic")) +
  theme_gray()

ppp2 = diabete_data %>%
  ggplot(aes(x = diabetes, y = HbA1c_level, fill = diabetes)) +
  geom_boxplot() +
  scale_fill_manual(values = c("#1f77b4", "#ff7f0e"), 
                    labels = c("Non-diabetic", "Diabetic")) +
  theme_gray()

grid.arrange(ppp1, ppp2, ncol = 2)
```

##### diabetes v.s. gender

```{r}
diabete_data %>%
  group_by(gender, diabetes) %>%
  summarise(count = n(), .groups = 'drop') %>%
  ggplot(aes(x = gender, y = count, fill = factor(diabetes))) +
  geom_col(width = 0.5) +  
  scale_fill_manual(values = c("#1f77b4", "#ff7f0e"), 
                    labels = c("Non-diabetic", "Diabetic")) +
  labs(title = "Diabetes Status Across Smoking Histories",
       x = "gender",
       y = "Count",
       fill = "Diabetes Status") +
  theme_gray()
```

##### diabetes v.s. smoking history

```{r}
diabete_data %>%
  group_by(smoking_history, diabetes) %>%
  summarise(count = n(), .groups = 'drop') %>%
  ggplot(aes(x = smoking_history, y = count, fill = factor(diabetes))) +
  geom_col(width = 0.5) +  
  scale_fill_manual(values = c("#1f77b4", "#ff7f0e"), 
                    labels = c("Non-diabetic", "Diabetic")) +
  labs(title = "Diabetes Status Across Smoking Histories",
       x = "Smoking History",
       y = "Count",
       fill = "Diabetes Status") +
  theme_gray()
```

#### Setting up model

```{r}
diabete_data_new = diabete_data %>%
  filter(gender != "Other")

diabete_data_new %>%
  group_by(gender) %>%
  summarise(count = n())

diabete_data_new$gender = factor(diabete_data_new$gender, 
                                    level = c("Female", "Male"))
```

##### Split dataset

```{r}
dia_split = diabete_data_new %>% initial_split(prop = 0.8, strata = diabetes)

dia_train = dia_split %>% training()
dia_test = dia_split %>% testing()

dim(dia_train)
dim(dia_test)
```

##### Create recipe

```{r}
dia_rec = 
  recipe(diabetes ~ ., data = dia_train) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_center(all_predictors()) %>%
  step_scale(all_predictors())

dia_rec %>%
  prep() %>%
  juice() %>%
  names()
```

##### Resampling with k fold cross-validation

```{r}
folds <- vfold_cv(dia_train, v = 10)
folds
```

#### Build up models 

##### Log-regression

```{r}
log_reg <- logistic_reg() %>% 
  set_engine("glm") %>% 
  set_mode("classification")

log_wflow <- workflow() %>%
  add_model(log_reg) %>%
  add_recipe(dia_rec)
```

##### LDA

```{r}
library(discrim)

lda_mod <- discrim_linear() %>% 
  set_mode("classification") %>% 
  set_engine("MASS")

lda_wflow <- workflow() %>%
  add_model(lda_mod) %>%
  add_recipe(dia_rec)
```

##### Naive Bayes

```{r}
library(klaR)

nb_mod <- naive_Bayes() %>% 
  set_mode("classification") %>% 
  set_engine("klaR") %>% 
  set_args(usekernel = FALSE) 

nb_wflow <- workflow() %>% 
  add_model(nb_mod) %>% 
  add_recipe(dia_rec)
```

##### random forest

```{r}
library(ranger)

rf_mod <- rand_forest(mtry = NULL, min_n = NULL, trees = 1000) %>%
  set_mode("classification") %>% 
  set_engine("ranger")

rf_wflow <- workflow() %>%
  add_model(rf_mod) %>%
  add_recipe(dia_rec)
```

#### Model evaluation

##### Log-regression

```{r}
# log-regression
log_result <- log_wflow %>%
  fit_resamples(
    resamples = folds, 
    metrics = metric_set(recall, precision, f_meas, accuracy, kap, roc_auc, sens, spec),
    control = control_resamples(save_pred = TRUE))

log_pred <- 
  log_result %>%
  collect_predictions()

log_pred %>%
  group_by(id) |># id contains our folds
  roc_curve(diabetes, .pred_0) |>
  autoplot()
```

##### LDA

```{r}
# LDA-regression
lda_result <- lda_wflow %>%
  fit_resamples(
    resamples = folds, 
    metrics = metric_set(recall, precision, f_meas, accuracy, kap, roc_auc, sens, spec),
    control = control_resamples(save_pred = TRUE))

lda_pred <- 
  lda_result %>%
  collect_predictions()

lda_pred %>%
  group_by(id) |># id contains our folds
  roc_curve(diabetes, .pred_0) |>
  autoplot()
```

##### Naive Bayes

```{r, message = FALSE}
# NB-regression
nb_result <- nb_wflow %>%
  fit_resamples(
    resamples = folds, 
    metrics = metric_set(recall, precision, f_meas, accuracy, kap, roc_auc, sens, spec),
    control = control_resamples(save_pred = TRUE))

nb_pred <- 
  nb_result %>%
  collect_predictions()

nb_pred %>%
  group_by(id) |># id contains our folds
  roc_curve(diabetes, .pred_0) |>
  autoplot()
```

##### Random Forest

```{r}
# LDA-regression
rf_result <- rf_wflow %>%
  fit_resamples(
    resamples = folds, 
    metrics = metric_set(recall, precision, f_meas, accuracy, kap, roc_auc, sens, spec),
    control = control_resamples(save_pred = TRUE))

rf_pred <- 
  lda_result %>%
  collect_predictions()

rf_pred %>%
  group_by(id) |># id contains our folds
  roc_curve(diabetes, .pred_0) |>
  autoplot()
```

#### Model selection by 

```{r}
log_metrics <- 
  log_result %>%
  collect_metrics() %>%
  mutate(model = "Logistic Regression")

lda_metrics <- 
  lda_result |>
  collect_metrics() %>%
  mutate(model = "Linear Discrimant Analysis") # add the name of the model to every row

nb_metrics <- 
  nb_result |>
  collect_metrics() %>%
  mutate(model = "Naive Bayes") # add the name of the model to every row

rf_metrics <- 
  rf_result |>
  collect_metrics() %>%
  mutate(model = "Random Forest")

model_compare <- bind_rows(log_metrics,
                           lda_metrics,
                           nb_metrics,
                           rf_metrics)

model_comp <- 
  model_compare %>%
  select(model, .metric, mean, std_err) %>%
  pivot_wider(names_from = .metric, values_from = c(mean, std_err))
```

```{r}
# show mean area under the ROC curve per model
model_comp |>
  arrange(mean_roc_auc) |>
  mutate(model = fct_reorder(model, mean_roc_auc)) %>%
  ggplot(aes(model, mean_roc_auc, fill=model)) +
  geom_col(width = 0.6) +
  coord_flip() +
  scale_fill_brewer(palette = "Blues") + 
     geom_text(
     size = 3,
     aes(label = round(mean_roc_auc, 2), y = mean_roc_auc + 0.08),
     vjust = 1
  )
```

#### Putting the model to the test dataset

```{r}
augment(rf_fit, new_data = dia_test) %>%
  roc_curve(diabetes, .pred_0) %>%
  autoplot()
```




